Based on polling as of 27 October 2024, Harris expects to win on xxx%, leads Trump by xx% in 2024 US election

### 2024-10-30 paper/paper.qmd
Change on data format

# Model

We conducted a regression analysis to predict support for Kamala Harris
in the 2024 U.S. presidential election. Our analysis includes two
models: an **unweighted linear model** and a **weighted linear model**.
These models predict Harris’s support as a continuous outcome across
different states, using aggregated polling data with a focus on
adjusting for poll quality, sample size, recency, and potential
partisanship biases. This setup aligns with the New York Times
methodology for election polling, which assigns weights based on poll
quality, sample size, and recency of the poll to reflect an accurate
picture of voter sentiment [@nytimes].

@@

We implemented our linear regression models using the `lm()` function in
R. The unweighted model provides a baseline with equal treatment for all
polls, while the weighted model adjusts each poll observation based on
combined factors, including recency, sample size, and poll quality. The
model specifications are as follows:

#### Unweighted Linear Model

The unweighted model serves as a baseline, treating all polls equally
without incorporating adjustments for recency, sample size, or poll
quality. The model is defined as:

$$
\text{Support_Harris}_i = \beta_0 + \beta_1 \cdot \text{National_Poll}_i + \epsilon_i
$$

This unweighted model provides a baseline estimate of Harris’s support
across different polls without considering variations in poll recency,
sample size, or quality. It allows us to understand Harris’s support
level with a simplified assumption that all polls are of equal
relevance.

#### Weighted Linear Model

To enhance the accuracy of our predictions, the weighted model
incorporates adjustments for recency, sample size, and pollster quality.
This model is defined as:

$$
\text{Support_Harris}_i = \beta_0 + \beta_1 \cdot \text{National_Poll}_i + \beta_2 \cdot \text{Recency_Weight}_i + \beta_3 \cdot \text{Sample_Size_Weight}_i + \beta_4 \cdot \text{Numeric_Grade}_i + \epsilon_i
$$

The weighted model applies these additional factors to account for
variations in data quality and recency, offering a refined view of
Harris’s support.

## Model Justification

In developing our model to predict support for Kamala Harris, we
included adjustments based on national polling, recency, sample size,
and pollster grade. These features were selected to mitigate common
biases in polling data and to ensure our model reflects up-to-date and
high-quality information. The decision to incorporate recency as a
predictor is motivated by the observation that voter sentiment can shift
rapidly in response to events, such as debates or political
controversies, especially as the election date approaches (The New York
Times, 2024). Sample size was also used as a weight, as larger polls
tend to reduce sampling variability up to a reasonable threshold, set at
2,300 responses.

Pollster quality, represented through numeric grades, accounts for
systematic differences in methodology, professionalism, and past
accuracy. Pollsters with a strong historical performance are given more
weight, as they are more likely to produce unbiased and reliable
results. The model structure, however, is based on linear regression,
meaning it assumes an additive relationship between these predictors and
voter support. This simplification, while computationally efficient and
interpretable, may not capture more complex interactions between
factors, such as shifts in demographic or regional support.

Our model was implemented in R using the `lm()` function for the
unweighted model and a weighted version for the adjusted model.
Validation was performed using out-of-sample testing and RMSE
calculations to assess accuracy. Sensitivity analyses were also
conducted, examining how the model performed across different subsets of
data.

The model has several limitations, including the linearity assumption,
which may not fully capture the dynamic nature of voter support. Its
effectiveness relies on the assumption that the weights (recency, sample
size, pollster grade) adequately mitigate biases in polling data.
Additionally, it does not account for third-party candidates or
undecided voters, which could affect the election outcome. While
Bayesian hierarchical models were considered for their capability to
handle multi-level data, they were ultimately set aside for a simpler,
more interpretable linear model due to the analysis's scope and
computational constraints. Future research should explore these more
complex methods if time and resources permit.

# Results

Our results are summarized in @tbl-modelresults.

@@

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/model_unweighted.rds"))
```

>>>>

## Model

We conducted a regression analysis to predict support for Kamala Harris
and Donald Trump in the 2024 U.S. presidential election, including
models for Trump for comparison. Our analysis includes four models:
**unweighted linear models** and **weighted linear models** for both
candidates. These models predict each candidate’s support as a
continuous outcome across different states, using **pollster**,
**national poll**, and **population** as key predictors. The weighted
models additionally incorporate adjustments for recency, sample size,
and poll quality to account for potential biases. This setup aligns with
the New York Times methodology for election polling, which assigns
weights based on poll quality, sample size, and recency to better
reflect voter sentiment [@nytimes].

@@

We implemented our linear regression models using the `lm()` function in
R.

### Mathematical Expressions

Both the unweighted and weighted models share the same mathematical
expression format:

#### Unweighted Model

The unweighted model for Harris provides a baseline by treating all
polls equally without adjustments for recency, sample size, or poll
quality:

$$
\text{Support\_Harris}_i = \beta_0 + \beta_1 \cdot \text{National\_Poll}_i + \beta_2 \cdot \text{Pollster}_i + \beta_3 \cdot \text{Population}_i + \epsilon_i
$$

```{r}
model_harris_unweighted <- lm(pct ~ national_poll + pollster + population, data = harris_data)
```

#### Weighted Model

The weighted model for Harris incorporates adjustments for recency,
sample size, and pollster quality:

$$
\text{Support\_Harris}_i = \beta_0 + \beta_1 \cdot \text{National\_Poll}_i + \beta_2 \cdot \text{Pollster}_i + \beta_3 \cdot \text{Population}_i + \epsilon_i
$$

```{r}
model_harris_weighted <- lm(pct ~ national_poll + pollster + population, data = harris_data, weights = combined_weight)
```

### Coefficient Explanations

-   $\beta_0$: Intercept of the model, representing the expected support
    when all predictors are zero.
-   $\beta_1$: Coefficient for the national poll, indicating how much
    support changes with a one-unit increase in the national poll
    percentage.
-   $\beta_2$: Coefficient for the pollster, reflecting the impact of
    different polling organizations on support.
-   $\beta_3$: Coefficient for the population, accounting for the
    influence of the demographic context on support levels.

### Model Justification

1.  **Use of Weighted Model**: The weighted model is essential for
    accurately reflecting voter sentiment. By incorporating weights, we
    adjust for the reliability of the polling data, ensuring that more
    credible polls have a greater influence on the estimates.

2.  **Calculation of Variables**:

    -   **Combined Weight**: The combined weight is calculated as
        follows: $$
        \text{combined\_weight} = \text{recency\_weight} \times \text{sample\_size\_weight} \times \text{poll\_frequency\_weight} \times \text{pollster\_quality\_weight}
        $$

    -   **Recency Weight**: This weight uses an exponential decay
        function: $$
        \text{recency\_weight} = \exp(-\text{Recency}_i \cdot 0.1)
        $$ This reflects the diminishing influence of older polls, with
        $\lambda = 0.1$ for the exponential decay.

    -   **Sample Size Weight**: This weight adjusts the significance of
        each poll based on the number of respondents, capping the
        weights at a maximum of 2,300 responses to reflect the
        reliability of larger sample sizes [@nyt_polling_averages].

    -   **Poll Frequency Weight**: This weight considers how often a
        pollster conducts polls, with higher weights assigned to
        pollsters with a greater number of recent surveys.

    -   **Pollster Quality Weight**: Based on the historical performance
        of the pollster, this weight emphasizes the reliability of their
        polling methods.

### Alternative Models

While this analysis employs linear regression with weighted least
squares, alternative methods such as Bayesian modeling could provide
additional insights. However, we opted for the weighted linear
regression approach to maintain interpretability and ease of
implementation given the scope and resources of this analysis. The
Bayesian approach was not chosen primarily due to the complexity
involved in defining priors and the additional computational
requirements.

### Weighted Least Squares Estimation

In our analysis, we utilize weighted least squares estimation to account
for the varying quality of polling data. The weights are crucial in the
estimation process, leading to the following expression for the
estimates of the coefficients ($\hat{\beta}$):

$$
\hat{\beta} = (X^T W X)^{-1} X^T W y
$$

Where: - $X$ is the design matrix of predictors. - $W$ is the diagonal
matrix of weights, which incorporates factors such as recency, sample
size, and pollster quality to enhance the reliability of our estimates.

### Importance of Combined Weights

By incorporating these combined weights, the weighted model offers a
more nuanced estimation of voter support for both Harris and Trump.
While the mathematical expression for the models remains consistent, the
weighted models effectively adjust for variations in data quality and
recency, leading to more accurate predictions of voter sentiment.

# Result

@@

#### Load Required Libraries ####
library(dplyr)
library(ggplot2)
library(lubridate)
harris_data <- harris_data %>%
  mutate(
    recency_weight = exp(-recency * 0.1),
    combined_weight = recency_weight * sample_size_weight * poll_frequency_weight * pollster_quality_weight,
    predicted_support = predict(model_harris_weighted, newdata = .),
    candidate = "Harris"
  )

trump_data <- trump_data %>%
  mutate(
    recency_weight = exp(-recency * 0.1),
    combined_weight = recency_weight * sample_size_weight * poll_frequency_weight * pollster_quality_weight,
    predicted_support = predict(model_trump_weighted, newdata = .),
    candidate = "Trump"
  )

# Combine both datasets
combined_data <- bind_rows(harris_data, trump_data)

#### Plot Support Trends by State with Reduced Date Labels ####
# Define states of interest
selected_states <- c("Pennsylvania", "Nevada", "North Carolina", "Wisconsin", "Michigan", "Georgia", "Arizona")

# Filter data for selected states
combined_data <- combined_data %>% filter(state %in% selected_states)

# Create the faceted plot with limited x-axis labels
ggplot(combined_data, aes(x = as.Date(end_date), y = predicted_support, color = candidate, size = combined_weight)) +
  geom_point(alpha = 0.6) +
  scale_color_manual(values = c("Harris" = "#1f77b4", "Trump" = "#d62728")) +  # Blue for Harris, Red for Trump
  scale_size_continuous(range = c(1, 4)) +  # Control point size range for readability
  scale_x_date(date_labels = "%b %d", date_breaks = "1 month") +  # Show only the first day of each month
  labs(title = "Support Trends for Harris and Trump Across States",
       x = "Date", y = "Support (%)", color = "Candidate") +
  facet_wrap(~ state, ncol = 2, scales = "free_y") +  # Facet by state with free y-scale
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "top",
    strip.text = element_text(size = 12, face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1, size = 8),  # Slight tilt for x-axis labels
    panel.grid.major = element_line(size = 0.2, linetype = "dotted", color = "gray80"),
    panel.grid.minor = element_blank()
  )
```
